---
title: "Getting Started with xtweetsR"
subtitle: "A complete workflow for collecting and analyzing X/Twitter data"
author: "Agustin Nieto"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with xtweetsR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Prerequisites

Before using **xtweetsR**, make sure you have:

- **R >= 4.2.0**
- **Node.js >= 18** (for the TypeScript sidecar)
- **Lightpanda** running and reachable at a CDP endpoint

Install the package from source:

```{r install, eval = FALSE}
devtools::install(".")
```

## Step 1: Check your environment

Always start with `x_doctor()` to verify that all dependencies are available:

```{r doctor, eval = FALSE}
library(xtweetsR)

x_doctor()
```

This runs eight independent checks:

| # | Check | What it verifies |
|---|-------|-----------------|
| 1 | R | Minimum version (4.2.0) |
| 2 | Node.js | Runtime availability |
| 3 | Sidecar | TypeScript compilation + ping response |
| 4 | Lightpanda connection | CDP connect request |
| 5 | CDP connection | Full browser session lifecycle |
| 6 | JavaScript evaluation | `Runtime.evaluate` works (`1 + 1 = 2`) |
| 7 | Network capture | CDP Network domain can be enabled |
| 8 | X navigation | Navigation to `https://x.com` |

A failure in one check does not prevent later checks from running. Checks 4-8 are skipped if the sidecar (check 3) is unavailable, since they all depend on it.

## Step 2: Create a session

A session starts the sidecar process and connects to Lightpanda:

```{r session, eval = FALSE}
session <- x_session()
print(session)
```

By default, the endpoint is resolved through this precedence:

1. `endpoint` argument in `x_session()`
2. `LPD_ENDPOINT` environment variable
3. Local default: `ws://127.0.0.1:21111`

You can also pass a custom endpoint or disable progress messages:

```{r session-custom, eval = FALSE}
session <- x_session(endpoint = "ws://192.168.1.100:21111", quiet = TRUE)
```

## Step 3: Search for posts

The core function is `x_search()`. It navigates to an X search results page, captures structured network responses, parses posts, deduplicates them, and returns a tibble:

```{r search, eval = FALSE}
results <- x_search(session, query = "rprogramming", limit = 10)
print(results)
```

Key parameters:

- **query** (required): Search term as a single string.
- **limit** (optional): Maximum number of posts to return.
- **scroll** (default `TRUE`): Whether to scroll the page to load more content.
- **max_scrolls** (default `5L`): Maximum scroll iterations when `scroll = TRUE`.

The function uses bounded repeated scrolling: after the initial extraction, the page is scrolled and new content is extracted in a loop. The loop stops when any of these conditions is met:

- The `limit` is reached.
- `max_scrolls` iterations have completed.
- No new data appears for two consecutive batches (stall detection).

## Step 4: Explore results with dplyr

`x_search()` returns a **tibble** with 26 columns, fully compatible with `dplyr`:

```{r dplyr, eval = FALSE}
library(dplyr)

# Most-liked posts
results %>%
  filter(like_count > 50) %>%
  arrange(desc(like_count))

# Posts by specific users
results %>%
  filter(username %in% c("rstudio", "hadleywickham")) %>%
  select(username, text, like_count, reply_count)

# Posts from a date range
results %>%
  filter(created_at >= "2025-01-01") %>%
  count(username, sort = TRUE)

# Inspect the collection provenance
attr(results, "rx_collection_provenance")
```

The returned tibble has these column groups:

- **Post identity**: `post_id`, `text`, `created_at`
- **Author**: `author_id`, `username`, `display_name`
- **Engagement metrics**: `reply_count`, `repost_count`, `like_count`, `quote_count`, `bookmark_count`, `view_count`
- **Relationships**: `conversation_id`, `is_reply`, `is_repost`, `is_quote`, `reply_to_post_id`, `quoted_post_id`
- **Entities**: `hashtags`, `mentions`, `urls` (list-columns)
- **Media**: `media_type`, `media_urls` (list-columns)
- **Provenance**: `collected_at`, `collection_query`, `collection_id`

## Step 5: Other collection functions

### User timeline

Collect posts from a specific user's timeline:

```{r user-posts, eval = FALSE}
rstudio_posts <- x_user_posts(session, username = "rstudio", limit = 20)
```

Parameters are the same as `x_search()`: `limit`, `scroll`, `max_scrolls`, `quiet`. You can also append a path segment like `"media"` or `"tweets_with_replies"`:

```{r user-path, eval = FALSE}
media_only <- x_user_posts(session, username = "rstudio", path = "media")
```

### Single post

Fetch a single post by its ID or URL:

```{r single-post, eval = FALSE}
post <- x_post(session, post_id = "1900000000000000001")
```

The `post_id` argument accepts both bare numeric IDs and full URLs.

## Step 6: Debug tools

Two functions help you inspect what's happening under the hood:

```{r debug, eval = FALSE}
# Inspect captured network traffic
network_df <- x_debug_network(session)
print(network_df)

# Inspect page DOM (full HTML)
full_html <- x_debug_dom(session)

# Or query specific elements
elements <- x_debug_dom(session, selector = "article")
print(elements)
```

## Step 7: Save your collection

Export results in multiple formats:

```{r export, eval = FALSE}
# JSONL -- always available
x_save(results, "collection.jsonl")

# Parquet -- requires the `arrow` package
x_save(results, "collection.parquet")

# DuckDB -- requires the `duckdb` package
x_save(results, "collection.duckdb")
```

JSONL is an append-only format ideal for incremental collection. Parquet and DuckDB support large collections that need SQL querying.

## Step 8: Resume and checkpoints

For long collections, save state and resume later:

```{r resume, eval = FALSE}
# First run -- saves checkpoint
results1 <- x_search(
  session,
  query = "rstats",
  limit = 100,
  checkpoint_path = "rstats.checkpoint.json",
  jsonl_path = "rstats.jsonl"
)

# Resume -- skips already-seen posts
results2 <- x_search(
  session,
  query = "rstats",
  limit = 100,
  resume = TRUE,
  checkpoint_path = "rstats.checkpoint.json",
  jsonl_path = "rstats.jsonl"
)
```

The checkpoint stores `collection_id`, `seen_post_ids`, `last_cursor`, and `records_collected`. When you resume, already-collected posts are not duplicated.

## Step 9: Close the session

Always close your session when finished:

```{r close, eval = FALSE}
x_close(session)
```

This stops the sidecar process, closes the CDP connection, and frees memory. Repeated calls are safe -- they return invisibly without error. A finalizer also runs automatically if your R session exits without calling `x_close()`.

## Summary

This vignette covered the complete workflow:

1. **Environment check** -- `x_doctor()`
2. **Session creation** -- `x_session()`
3. **Search** -- `x_search()` with scrolling and deduplication
4. **Explore** -- `dplyr` on the returned tibble
5. **Other collectors** -- `x_user_posts()`, `x_post()`
6. **Debug** -- `x_debug_network()`, `x_debug_dom()`
7. **Export** -- `x_save()` (JSONL, Parquet, DuckDB)
8. **Resume** -- checkpoint-based collection continuity
9. **Cleanup** -- `x_close()`

All collection functions share the same architecture: network-first extraction via CDP, canonical normalization, deduplication, and provenance tracking. This means you can combine results from different queries and be confident the schema is consistent.
