---
title: "Large Collections"
subtitle: "Managing limits, checkpoints, resume, and export for large-scale collection"
author: "Agustin Nieto"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Large Collections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

When collecting data at scale from X/Twitter, three concerns dominate:

1. **Memory** -- downloading thousands of posts in one call can exhaust available RAM.
2. **Resilience** -- network failures, Lightpanda crashes, or X rate-limits can interrupt a long collection.
3. **Storage** -- once collected, the data needs to be saved in a format suitable for analysis (JSONL for incremental pipelines, Parquet for columnar analytics, DuckDB for SQL queries).

This vignette covers how to handle each concern with **xtweetsR**.

## The `limit` parameter

Every collection function (`x_search()`, `x_user_posts()`) accepts a `limit` parameter:

```r
results <- x_search(session, query = "rstats", limit = 100)
```

`limit` caps the total number of **unique** posts returned. The function stops scrolling and returns early once the limit is reached. If no limit is specified (`NULL`, the default), the function uses bounded scrolling (`max_scrolls` iterations with stall detection) but does not impose a hard cap on the result count.

**Recommended practice:** Always set `limit` for production collection. A typical research query might use `limit = 500` to `limit = 5000`. Larger limits require more memory and more scrolling iterations.

```{r, eval = FALSE}
# Collect up to 200 posts about a topic.
results <- x_search(session, query = "machine learning", limit = 200)

# Collect all available posts (use with caution).
all_posts <- x_search(session, query = "rstats")
```

## Checkpoints

A checkpoint is a JSON file that stores the collection state at a point in time:

```
{
  "collection_id": "550e8400-e29b-41d4-a716-446655440000",
  "query": "rstats",
  "seen_post_ids": ["1234567890123456789", "1234567890123456790", ...],
  "last_cursor": "cursor-batch-3",
  "last_post_id": "1234567890123456999",
  "records_collected": 147
}
```

The checkpoint is written by `.rx_checkpoint_write()` at the end of a search when `resume = TRUE`. You can specify where it is stored:

```{r, eval = FALSE}
# The checkpoint defaults to a temp file, but you should specify it explicitly
# so you can resume later.
results <- x_search(
  session,
  query = "rstats",
  limit = 1000,
  checkpoint_path = "collections/rstats.checkpoint.json",
  jsonl_path = "collections/rstats.jsonl"
)
```

Checkpoint fields:

| Field | Type | Description |
|-------|------|-------------|
| `collection_id` | character | UUID that identifies this collection run |
| `query` | character | The search query string |
| `seen_post_ids` | character vector | All post IDs collected so far |
| `last_cursor` | character | Cursor from the last network response |
| `last_post_id` | character | ID of the last post collected |
| `records_collected` | integer | Count of unique posts collected |

## Resume workflow

Resume allows you to continue a collection from where it left off. The workflow is:

### First run

```{r, eval = FALSE}
# Collect the first batch. The checkpoint is written at the end.
batch1 <- x_search(
  session,
  query = "rstats",
  limit = 500,
  checkpoint_path = "collections/rstats.checkpoint.json",
  jsonl_path = "collections/rstats.jsonl"
)

# batch1 contains up to 500 posts. The checkpoint saves the state.
print(attr(batch1, "rx_collection_provenance"))
```

### Resume

```{r, eval = FALSE}
# Resume continues from the checkpoint. Already-collected posts are skipped.
batch2 <- x_search(
  session,
  query = "rstats",
  limit = 500,
  resume = TRUE,
  checkpoint_path = "collections/rstats.checkpoint.json",
  jsonl_path = "collections/rstats.jsonl"
)

# batch2 contains only NEW posts not in the checkpoint.
# The total unique posts = nrow(batch1) + nrow(batch2), with no duplicates.
```

### How resume works

1. **Checkpoint read:** `.rx_checkpoint_read()` loads the saved state from `checkpoint_path`.
2. **State pre-population:** `.rx_scroll_state_new(seen_post_ids = checkpoint$seen_post_ids)` initializes the scroll state with previously seen IDs.
3. **Deduplication:** New posts are deduplicated against `seen_post_ids` before being added to the result.
4. **Checkpoint write:** After the search completes, `.rx_checkpoint_write()` overwrites the checkpoint with the updated state.

The `collection_id` is preserved across resume calls so all posts from a multi-run collection share the same identifier.

### Resume without a checkpoint

If `resume = TRUE` but no checkpoint file exists, the function behaves normally (no resume, no error). This makes it safe to always set `resume = TRUE` in scripts:

```{r, eval = FALSE}
# Safe to call even on the first run -- behaves like resume = FALSE
# when no checkpoint exists.
results <- x_search(
  session,
  query = "rstats",
  limit = 200,
  resume = TRUE,
  checkpoint_path = "collections/rstats.checkpoint.json",
  jsonl_path = "collections/rstats.jsonl"
)
```

### JSONL incremental persistence

While `checkpoint_path` stores collection *state*, `jsonl_path` stores the actual post data in append-only JSONL format. Each call to `x_search()` with `resume = TRUE` appends new posts to the JSONL file rather than overwriting it.

JSONL (JSON Lines) is ideal for incremental pipelines because:
- Each line is a self-contained JSON object (one post per line).
- Appending is safe -- no need to read the entire file.
- The file can be read back at any time with `.rx_jsonl_read()`.
- It works as an intermediate format between collection and final storage.

```{r, eval = FALSE}
# Read a JSONL file back into a tibble.
posts <- .rx_jsonl_read("collections/rstats.jsonl")
print(nrow(posts))  # Total posts across all collection runs
```

## Export formats

The `x_save()` function writes a post tibble to disk. The format is determined by the file extension:

```{r, eval = FALSE}
# JSONL -- always available, append-friendly
x_save(results, "results.jsonl")

# Parquet -- requires the `arrow` package
x_save(results, "results.parquet")

# DuckDB -- requires the `duckdb` package
x_save(results, "results.duckdb")
```

### JSONL

Always available (uses `jsonlite`). Each row becomes one JSON object on a line.

**Best for:** Incremental collection pipelines, data transfer between systems, human-readable inspection.

**Trade-offs:**
- Pros: Simple, append-friendly, no external dependencies.
- Cons: No columnar compression, slower for large-file scanning, no SQL support.

### Parquet

Uses the `arrow` package (`Suggests`). Columnar binary format with compression.

**Best for:** Analytics pipelines, large datasets, interoperability with Spark/Pandas.

**Trade-offs:**
- Pros: Columnar compression (smaller files), fast column scans, widely supported.
- Cons: Requires `arrow` (falls back to JSONL with a warning if missing), not append-friendly.

### DuckDB

Uses the `duckdb` package (`Suggests`). Creates a database file with a single `posts` table.

**Best for:** Interactive SQL analysis, ad-hoc queries, local data warehouse.

**Trade-offs:**
- Pros: Full SQL support, fast queries, can query files directly.
- Cons: Requires `duckdb` (falls back to JSONL with a warning if missing), not a simple file format.

### Fallback behavior

If the required package (`arrow` or `duckdb`) is not installed, `x_save()` falls back to JSONL with a warning:

```
Warning: The 'arrow' package is not installed.
Falling back to JSONL at 'results.jsonl'.
Install arrow for native Parquet support.
```

The fallback uses a `.jsonl` extension so the user knows the format changed.

## Complete large-collection pattern

Here is a complete example that combines all the concepts:

```{r, eval = FALSE}
library(xtweetsR)

# 1. Start session.
session <- x_session()

# 2. Define collection parameters.
query      <- "rstats"
limit      <- 1000
max_scrolls <- 20L
base_dir   <- "collections/rstats"

# Create the output directory if needed.
dir.create(base_dir, showWarnings = FALSE, recursive = TRUE)

checkpoint_path <- file.path(base_dir, "checkpoint.json")
jsonl_path      <- file.path(base_dir, "posts.jsonl")
parquet_path    <- file.path(base_dir, "posts.parquet")

# 3. Check for an existing checkpoint.
existing_checkpoint <- .rx_checkpoint_read(checkpoint_path)
resume <- !is.null(existing_checkpoint)

# 4. Run the search (resume if checkpoint exists).
results <- x_search(
  session,
  query = query,
  limit = limit,
  max_scrolls = max_scrolls,
  resume = resume,
  checkpoint_path = checkpoint_path,
  jsonl_path = jsonl_path
)

# 5. Show collection provenance.
cat("Collection metadata:\n")
print(attr(results, "rx_collection_provenance"))

# 6. Save to Parquet for analytics.
x_save(results, parquet_path)

# 7. Close session.
x_close(session)
```

## Troubleshooting

### Resume not picking up new posts

Check that `checkpoint_path` and `jsonl_path` are the same on every call. If the path changes between calls, resume will not work.

### Checkpoint file not written

The checkpoint is only written when `resume = TRUE`. If you want to save state without resuming, call `.rx_checkpoint_write()` manually:

```{r, eval = FALSE}
# Save state manually (not tied to a search run).
# The scroll state object must be passed from a running search.
```

### JSONL file grows very large

JSONL is append-only. If you run multiple search calls with the same `jsonl_path`, the file accumulates all posts. Use `x_save()` with `.jsonl` extension (append=FALSE by default) to create a fresh snapshot:

```{r, eval = FALSE}
# Create a clean snapshot from accumulated JSONL.
accumulated <- .rx_jsonl_read("collections/rstats.jsonl")
x_save(accumulated, "collections/rstats-snapshot.jsonl")
```

### Parquet vs JSONL for large collections

For collections over 100K posts:
- Use Parquet for storage (smaller files, faster reads).
- Use JSONL for incremental collection (append-friendly).
- Convert JSONL to Parquet after collection using `x_save(.rx_jsonl_read(path), "output.parquet")`.

## Function reference

### `x_search()`

```r
x_search(
  session,
  query,
  limit = NULL,
  scroll = TRUE,
  max_scrolls = 5L,
  resume = FALSE,
  checkpoint_path = NULL,
  jsonl_path = NULL,
  quiet = FALSE
)
```

| Parameter | Type | Default | Description |
|-----------|------|---------|-------------|
| `session` | `xtweetsR_session` | (required) | Browser session from `x_session()` |
| `query` | character | (required) | Search query string |
| `limit` | integer or NULL | NULL | Maximum posts to return |
| `scroll` | logical | TRUE | Enable scroll-based collection |
| `max_scrolls` | integer | 5L | Max scroll iterations |
| `resume` | logical | FALSE | Resume from checkpoint |
| `checkpoint_path` | character or NULL | NULL | Path to checkpoint JSON file |
| `jsonl_path` | character or NULL | NULL | Path to JSONL incremental file |
| `quiet` | logical | FALSE | Suppress progress messages |

### `x_save()`

```r
x_save(posts, path)
```

| Parameter | Type | Description |
|-----------|------|-------------|
| `posts` | tibble | Post tibble from `x_search()` or similar |
| `path` | character | Output file path (extension determines format) |

Supported extensions: `.jsonl`, `.parquet`, `.duckdb`.

### Persistence helpers

```r
# Read JSONL file back into a tibble.
.rx_jsonl_read(path)

# Write tibble to JSONL.
.rx_jsonl_write(path, posts, append = TRUE)

# Read checkpoint file.
.rx_checkpoint_read(path)

# Write checkpoint file.
.rx_checkpoint_write(path, checkpoint)
```
